---
title: "Illustrating the Use of EcoTourImpact"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tourism_analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,       # Show code output by default
  warning = FALSE,   # Suppress warnings in output
  message = FALSE    # Suppress messages in output
)
```


# Introduction

The `EcoTourImpact` package provides tools to analyze the contribution of tourism to GDP across various countries and years. The data includes metrics on tourism's percentage share of total GDP, which allows users to understand the economic impact of tourism globally.

## Package Installation

```{r}
# Specify a CRAN mirror and install `remotes`
install.packages("remotes", repos = "https://cloud.r-project.org/")
```

You can install the package from GitHub:

```{r}
# Install TourismGDP
remotes::install_github("ETC5523-2024/assignment-4-packages-and-shiny-apps-DubeyArushi19")

# Load the package
library(EcoTourImpact)
```

## Loading Data and Libraries

After installing the package, load the necessary libraries and the dataset provided by EcoTourImpact

```{r echo=FALSE}
#Loading the libraries

library(dplyr)
library(ggplot2)
library(EcoTourImpact)

#Reading the dataset

tourism <- read_csv("data/tourism-gdp-proportion-of-total-gdp.csv")
```

## Exploring Tourism GDP Data

The tourism dataset contains data on the percentage of GDP generated by tourism across various countries and years. Let’s take a closer look at the structure of the dataset:

```{r}
# Display the first few rows of the dataset
head(tourism)
```

### Data Description

- **Entity**: The name of the country.
- **Code**: The country code.
- **Year**: The year of the observation.
- **GDP from tourism as a share of total GDP**: The percentage of the country's GDP generated by tourism.



## Example: Analyzing Tourism's Contribution to GDP
Let’s explore tourism’s impact on a specific country over time and compare it with other countries. We’ll use this dataset to visualize trends in tourism's contribution to GDP.

### Data Preparation

- Filter data for a single country "Australia" 

```{r}
# Filter data for a single country 
australia_data <- tourism %>%
  filter(Entity == "Australia", Year >= 2008, Year <= 2022)

# Display filtered data
head(australia_data)
```


### Visualization: Tourism GDP Over Time

Using ggplot2, we can plot tourism's GDP contribution over time to observe any trends.

```{r}
#| label: fig-1
#| fig-cap: The graph demonstrate the trends in the contribution of tourism to Australia's GDP from 2008 to 2022.

#Create a line chart

ggplot(australia_data, aes(x = Year, y = `GDP from tourism as a share of total GDP`)) +
  geom_line(color = "blue", size = 1) + 
  geom_point(color = "red", size = 2) +     
  labs(
    title = "Proportion of GDP from Tourism in Australia (2008-2022)",
    x = "Year",
    y = "GDP from Tourism as a Share of Total GDP (%)"
  ) +
  theme_minimal() 
```

This line plot shows the changes in tourism’s contribution to Australia’s GDP over time, providing insight into how reliant Australia’s economy is on tourism.

### Comparing Multiple Countries
Let’s compare tourism’s GDP contribution in several countries over a defined period. This example demonstrates how to prepare data for multiple countries and visualize them on the same plot.

```{r}
# Filter data for multiple countries
countries_data <- tourism %>%
  filter(Entity %in% c("Australia", "United States", "France", "Japan"),
         Year >= 2008, Year <= 2020)

# Plot tourism GDP contribution for multiple countries
ggplot(countries_data, aes(x = Year, y = `GDP from tourism as a share of total GDP`, color = Entity)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Comparison of Tourism GDP Contribution Across Countries",
    x = "Year",
    y = "Tourism GDP as % of Total GDP",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
  )

```
This plot compares the tourism GDP contributions for Australia, the United States, France, and Japan, allowing us to observe differences in tourism’s economic impact across countries.

## Interpretation of the Plot

- **Australia**: Observe how Australia’s tourism GDP contribution changes over time, with peaks and valleys that may reflect economic events, policy changes, or global tourism trends.

- **United States, France, and Japan**: Each country's line indicates how heavily it relies on tourism relative to its total GDP. Notable differences may highlight varying levels of tourism dependency or resilience to economic shifts.
This side-by-side visualization provides an intuitive comparison of tourism’s economic role in different countries.

## Conclusion

The **EcoTourImpact** package provides preprocessed datasets and functions that facilitate exploration and visualization of tourism’s impact on GDP. This vignette demonstrated how to load the data, filter it for analysis, and visualize tourism's GDP contribution over time and across countries.

By using **EcoTourImpact**, users can gain insights into tourism's economic role globally, and explore trends and patterns in tourism’s impact on national economies.
