---
title: "Illustrating the Use of EcoTourImpact"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tourism_analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,       # Show code output by default
  warning = FALSE,   # Suppress warnings in output
  message = FALSE    # Suppress messages in output
)
```


# Introduction

The `EcoTourImpact` package provides tools to analyze the contribution of tourism to GDP across various countries and years. The data includes metrics on tourism's percentage share of total GDP, which allows users to understand the economic impact of tourism globally.

You can install the package from GitHub:


## Loading Data and Libraries

After installing the package, load the necessary libraries and the dataset provided by EcoTourImpact

```{r echo=FALSE}
#Loading the libraries

library(dplyr)
library(ggplot2)
library(EcoTourImpact)
```

```{r}
#Loading the dataset

load("D:/MONASH/ETC 5523 CD/Assignment 4/assignment-4-packages-and-shiny-apps-DubeyArushi19/data/tourism.rda")
```


### Data Description

- **Entity**: The name of the country.
- **Code**: The country code.
- **Year**: The year of the observation.
- **GDP from tourism as a share of total GDP**: The percentage of the country's GDP generated by tourism.



## Example: Analyzing Tourism's Contribution to GDP
Let’s explore tourism’s impact on a specific country over time and compare it with other countries. We’ll use this dataset to visualize trends in tourism's contribution to GDP.

### Data Preparation

- Filter data for a single country "Australia" 

```{r}
# Filter data for a single country 
australia_data <- tourism %>%
  filter(Entity == "Australia", Year >= 2008, Year <= 2022)

# Display filtered data
head(australia_data)
```


### Visualization: Tourism GDP Over Time

Using ggplot2, we can plot tourism's GDP contribution over time to observe any trends.

```{r}
#| label: fig-1
#| fig-cap: The graph demonstrate the trends in the contribution of tourism to Australia's GDP from 2008 to 2022.

#Create a line chart

ggplot(australia_data, aes(x = Year, y = `GDP from tourism as a share of total GDP`)) +
  geom_line(color = "blue", size = 1) + 
  geom_point(color = "red", size = 2) +     
  labs(
    title = "Proportion of GDP from Tourism in Australia (2008-2022)",
    x = "Year",
    y = "GDP from Tourism as a Share of Total GDP (%)"
  ) +
  theme_minimal() 
```

This line plot shows the changes in tourism’s contribution to Australia’s GDP over time, providing insight into how reliant Australia’s economy is on tourism.

### Comparing Multiple Countries
Let’s compare tourism’s GDP contribution in several countries over a defined period. 

```{r}
# Filter data for multiple countries
countries_data <- tourism %>%
  filter(Entity %in% c("Australia", "United States", "France", "Japan"),
         Year >= 2008, Year <= 2020)

# Plot tourism GDP contribution for multiple countries
ggplot(countries_data, aes(x = Year, y = `GDP from tourism as a share of total GDP`, color = Entity)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Comparison of Tourism GDP Contribution Across Countries",
    x = "Year",
    y = "Tourism GDP as % of Total GDP",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
  )

```
This plot compares the tourism GDP contributions for Australia, the United States, France, and Japan, allowing us to observe differences in tourism’s economic impact across countries.

## Interpretation of the Plot

- **Australia**: Observe how Australia’s tourism GDP contribution changes over time, with peaks and valleys that may reflect economic events, policy changes, or global tourism trends.

- **United States, France, and Japan**: Each country's line indicates how heavily it relies on tourism relative to its total GDP. Notable differences may highlight varying levels of tourism dependency or resilience to economic shifts.
This side-by-side visualization provides an intuitive comparison of tourism’s economic role in different countries.

## Functions for additional information:

- 1.**Filter Tourism GDP Data by Country and Year Range**

```{r}
# Filter tourism data for Australia from 2008 to 2020
filtered_data <- filter_tourism_data(tourism, "Australia", c(2008, 2020))
head(filtered_data)

```

- 2.**Plot Tourism GDP Contribution Over Time**

```{r}
# Plot the tourism GDP contribution for Australia
australia_data <- filter_tourism_data(tourism, "Australia", c(2008, 2020))
plot_tourism_gdp(australia_data)

```

- 3.**Summarize Tourism GDP Contribution Statistics**

```{r}
# Summarize tourism GDP statistics for Australia
summary_stats <- summarize_tourism_stats(australia_data)
print(summary_stats)

```


## Conclusion

This vignette shows how EcoTourImpact provides tools for exploring tourism GDP data. With functions for filtering, visualization, and summary statistics, the package enables in-depth analysis and interactive exploration. The Shiny app further enhances this experience by offering an intuitive interface for users who prefer interactive data exploration.
